### 2024-08-11 download-football-data.qmd
20240810 - updates

```{r}
nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly_raw %>% 
  select(player_id, season, week, everything()) %>% 
  arrange(player_display_name, player_id, season, week)

@@

```{r}
nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal_raw %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)

@@

```{r}
# Reshape from wide to long format
nfl_actualStats_offense_weekly_long <- nfl_actualStats_offense_weekly %>% 
  tidyr::pivot_longer(
    cols = c(recent_team, opponent_team),
    names_to = "role",
    values_to = "team")

@@

# Merge with player info
player_age_offense <- dplyr::left_join(
  player_game_birthdate_gameday_offense,
  nfl_players %>% select(-birth_date, -season),
  by = c("player_id" = "gsis_id"))

player_age_defense <- dplyr::left_join(
  player_game_birthdate_gameday_defense,
  nfl_players %>% select(-birth_date, -season),
  by = c("player_id" = "gsis_id"))

player_age_kicking <- dplyr::left_join(
  player_game_birthdate_gameday_kicking,
  nfl_players %>% select(-birth_date, -season),
  by = c("player_id" = "gsis_id"))

@@

```{r}
# Merge player info with seasonal stats
player_stats_seasonal_offense <- dplyr::full_join(
  nfl_actualStats_offense_seasonal,
  nfl_players %>% select(-position, -position_group, -season),
  by = c("player_id" = "gsis_id")
)

player_stats_seasonal_defense <- dplyr::full_join(
  nfl_actualStats_defense_seasonal,
  nfl_players %>% select(-position, -position_group, -season),
  by = c("player_id" = "gsis_id")
)

player_stats_seasonal_kicking <- dplyr::full_join(
  nfl_actualStats_kicking_seasonal,
  nfl_players %>% select(-position, -position_group, -season),
  by = c("player_id" = "gsis_id")
)

>>>>

```{r}
nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly_raw %>% 
  rename(team = recent_team) %>% 
  select(player_id, season, week, everything()) %>% 
  arrange(player_display_name, player_id, season, week)

@@

```{r}
nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal_raw %>% 
  rename(team = recent_team) %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)

@@

```{r}
# Reshape from wide to long format
nfl_actualStats_offense_weekly_long <- nfl_actualStats_offense_weekly %>% 
  tidyr::pivot_longer(
    cols = c(team, opponent_team),
    names_to = "role",
    values_to = "team")

@@

# Merge with player info
player_age_offense <- dplyr::left_join(
  player_game_birthdate_gameday_offense,
  nfl_players %>% select(-birth_date, -season, -team_abbr, - team_seq),
  by = c("player_id" = "gsis_id"))

player_age_defense <- dplyr::left_join(
  player_game_birthdate_gameday_defense,
  nfl_players %>% select(-birth_date, -season, -team_abbr, - team_seq),
  by = c("player_id" = "gsis_id"))

player_age_kicking <- dplyr::left_join(
  player_game_birthdate_gameday_kicking,
  nfl_players %>% select(-birth_date, -season, -team_abbr, - team_seq),
  by = c("player_id" = "gsis_id"))

@@

```{r}
# Merge player info with seasonal stats
player_stats_seasonal_offense <- dplyr::full_join(
  nfl_actualStats_offense_seasonal,
  nfl_players %>% select(-position, -position_group, -season, -team_abbr, - team_seq),
  by = c("player_id" = "gsis_id")
)

player_stats_seasonal_defense <- dplyr::full_join(
  nfl_actualStats_defense_seasonal,
  nfl_players %>% select(-position, -position_group, -season, -team_abbr, - team_seq),
  by = c("player_id" = "gsis_id")
)

player_stats_seasonal_kicking <- dplyr::full_join(
  nfl_actualStats_kicking_seasonal,
  nfl_players %>% select(-position, -position_group, -season, -team_abbr, - team_seq),
  by = c("player_id" = "gsis_id")
)

### 2024-08-11 machine-learning.qmd
20240810 - updates

```{r}
library("LongituRF")
library("doParallel")
library("missForest")
library("powerjoin")
library("mice")
library("miceadds")
library("caret")
library("tidyverse")
```

@@

## Rename other variables to ensure common names
player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
  rename(team = team_abbr)

@@

seasonalData <- powerjoin::power_full_join(
  playerSeasonMerged,
  playerMerged,
  by = "gsis_id",
  conflict = coalesce_xy
) %>% 
  filter(!is.na(season)) %>% 
  select(gsis_id, season, player_display_name, position, recent_team, games, everything()) #-todo: deduplicate team from prior merged objects

@@

# check on duplicate id-by-season data (nfl_rosters, nfl_advancedStatsPFR_seasonal)

seasonalData %>% 
  group_by(gsis_id, season) %>% 
  filter(n() > 1) %>% 
  head()

@@

newData_seasonalQB <- left_join(
  newData_seasonalQB,
  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
  by = "gsis_id"
)

newData_seasonalRB <- left_join(
  newData_seasonalRB,
  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
  by = "gsis_id"
)

newData_seasonalWR <- left_join(
  newData_seasonalWR,
  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
  by = "gsis_id"
)

newData_seasonalTE <- left_join(
  newData_seasonalTE,
  newData_seasonal %>% select(gsis_id, player_display_name, recent_team),
  by = "gsis_id"
)

@@

>>>>

```{r}
library("doParallel")
library("missForest")
library("powerjoin")
library("caret")
library("tidyverse")
```

@@

## Rename other variables to ensure common names

@@

seasonalData <- powerjoin::power_full_join(
  playerSeasonMerged,
  playerMerged %>% select(-team, -team_abbr, -team_seq, -current_team_id),
  by = "gsis_id",
  conflict = coalesce_xy
) %>% 
  filter(!is.na(season)) %>% 
  select(gsis_id, season, player_display_name, position, team, games, everything())

@@

# Duplicate cases
seasonalData %>% 
  group_by(gsis_id, season) %>% 
  filter(n() > 1) %>% 
  head()

@@

newData_seasonalQB <- left_join(
  newData_seasonalQB,
  newData_seasonal %>% select(gsis_id, player_display_name, team),
  by = "gsis_id"
)

newData_seasonalRB <- left_join(
  newData_seasonalRB,
  newData_seasonal %>% select(gsis_id, player_display_name, team),
  by = "gsis_id"
)

newData_seasonalWR <- left_join(
  newData_seasonalWR,
  newData_seasonal %>% select(gsis_id, player_display_name, team),
  by = "gsis_id"
)

newData_seasonalTE <- left_join(
  newData_seasonalTE,
  newData_seasonal %>% select(gsis_id, player_display_name, team),
  by = "gsis_id"
)

@@

library("LongituRF")

### 2024-08-11 mixed-models.qmd
20240810 - updates

```{r}
fantasyPointsVector <- player_stats_seasonal_offense_subset$fantasy_points %>% 
    na.omit() %>% 
    as.vector()

player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasy_points
player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasy_points < 0] <- 0
player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasy_points <= 0] <- 0.01

@@

```{r}
plot(fit.norm)
```

```{r}
plot(fit.lognormal)
```

```{r}
plot(fit.gamma)
```

```{r}
plot(fit.exp)
```

```{r}
plot(fit.weibull)
```

>>>>

```{r}
player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros <- player_stats_seasonal_offense_subset$fantasyPointsPos <- player_stats_seasonal_offense_subset$fantasy_points
player_stats_seasonal_offense_subset$fantasyPointsPos[player_stats_seasonal_offense_subset$fantasy_points < 0] <- 0
player_stats_seasonal_offense_subset$fantasyPointsPosNoZeros[player_stats_seasonal_offense_subset$fantasy_points <= 0] <- 0.01

fantasyPointsVector <- player_stats_seasonal_offense_subset$fantasy_points %>% 
    na.omit() %>% 
    as.vector()

@@

```{r}
#| label: fig-normalDistribution
#| fig-cap: "Fit of normal distribution to fantasy points."

plot(fit.norm)
```

```{r}
#| label: fig-logNormalDistribution
#| fig-cap: "Fit of log normal distribution to fantasy points."

plot(fit.lognormal)
```

```{r}
#| label: fig-gammaDistribution
#| fig-cap: "Fit of gamma distribution to fantasy points."

plot(fit.gamma)
```

```{r}
#| label: fig-exponentialDistribution
#| fig-cap: "Fit of exponential distribution to fantasy points."

plot(fit.exp)
```

```{r}
#| label: fig-weibullDistribution
#| fig-cap: "Fit of Weibull distribution to fantasy points."

plot(fit.weibull)
```
